Convergence rate of MAP estimates \\ for the exponential family

### 2021-10-13 aistats.tex
Remove MAP=SMD subsection

\paragraph{Optimization}
We can rewrite~\eqref{eq:defMAP} as
\alignn{
\m_n = \m_{n-1}- \lr_n (\m_{n-1} - T_n)
\label{eq:mean_update}
}
where $\lr_n := \inv{n_0 + n}$. 
Now let $f_X(\nat) = -\log p(X \cond \nat)$ such that $\E[f_X] = f$.
If we further introduce $g_n(\nat) := \nabla\logpart(\nat) - T_n = \nabla f_{X_n}(\nat)$, then~\eqref{eq:mean_update} becomes
\alignn{
\nabla\conj(\hat \nat_{n}) 
= \nabla\conj(\hat \nat_{n-1}) - \lr_n g_n(\hat \nat_{n-1})
}
which we recognize as the update formula for stochastic mirror descent (SMD) on $f$ with mirror map $\nabla\logpart$ and step-size schedule $\lr_n$.
\emph{MAP is an instantiation of SMD.}

@@

\section{Optimization Perspective}
\label{sec:optimization}

We now show how MAP can be interpreted as stochastic mirror descent (SMD).
This means that \textbf{1)} we may obtain a convergence rate for MAP from an optimization analysis, and \textbf{2)} any insights gained from the MAP may inform further design and analysis of SMD.
We then review the assumptions of relative smoothness, useful to deal with non-smooth functions, before investigating 3 analysis of SMD with the MAP.
\rlp{mention that examples have so far been lacking for relative SMD.}

\subsection{MAP as Stochastic Mirror Descent}
\label{ssec:MAP=SMD}
By Bayes rule, the posterior can always be written iteratively $p(\nat | \mD_n) \propto p(X_n | \nat) p(\nat | \mD_{n-1})$.
Plugging in~\eqref{eq:joint_likelihood}, and introducing $f_n(\nat) = - \log p(X_{n} | \nat)$, we see that the MAP verifies
\begin{align*}
	\hat \nat_{n}
	= \argmin_\nat f_n(\nat) + (n_0 + n -1) \bregman(\nat; \hat \nat_{n-1})
    %\label{eq:SBPP}
\end{align*}
which is the update of a Stochastic Bregman Proximal Point update with step-size $\inv{n_0 + n - 1}$.
%Maybe skip this SBPP story altogether
Expanding $f_n(\nat) = \logpart(\nat) - T_n$ and  playing with the Bregman we get
\begin{align*}
		\hat \nat_{n}
	= \argmin_\nat \langle g_{n}(\hat \nat_{n-1}), \nat \rangle + (n_0 + n) \bregman(\nat; \hat \nat_{n-1})
    %\label{eq:SMD}
\end{align*}
where $g_{n}(\nat) := \nabla f_n(\nat) = \hat \m_{n-1} - T_n$.
We recognize the update of Stochastic Mirror Descent (a.k.a. Stochastic Bregman Gradient) with step-size $\lr_n = \inv{n_0 + n}$.
From the first order optimality condition, we see this is a stochastic gradient step in the dual $\hat \m_n = \hat \m_{n-1} - \lr_n g_n$.
This parallel means we could obtain a finite convergence rate for the  MAP from an analysis of these algorithms.
In particular, it could potentially apply to a wide range of distribution $\cD$ on $X$.
Note that this optimization perspective does not hold for the MLE %($n_0=0$)
as its implicit uniform prior may correspond to an initialization outside of $\Theta$.

\subsection{Relative Smoothness}
Mirror descent (MD)\footnote{
Also known as
Bregman (proximal) gradient, relative gradient descent or NoLips.
}
\citep{nemirovski1983problem,beck2003mirror}
%Nemirovski introduced the algorithm, and Beck established the connection with Bregman divergences, e.g. the argmin form.
and SMD 
\citep{nemirovski2009robust,ghadimi2012optimal} 
are typically encountered in non-smooth (online) optimization,
under bounded (or Lipschitz) gradient assumption on the objective $f$
and strong convexity assumption on the potential $\logpart$
\citep[Th.4.2 \& Th.6.3]{bubeck2015convex}.
In our case, none of these assumptions may hold. 
For instance $\logpart = -\log$ is neither smooth nor strongly convex.

>>>>

\paragraph{Optimization}
We can rewrite~\eqref{eq:defMAP} as
\alignn{
\m_n = \m_{n-1}- \lr_n (\m_{n-1} - T_n)
\label{eq:mean_update}
}
where $\lr_n := \inv{n_0 + n}$. 
Now let $f_X(\nat) = -\log p(X \cond \nat)$ such that $\E[f_X] = f$.
If we further introduce $g_n(\nat) := \nabla\logpart(\nat) - T_n = \nabla f_{X_n}(\nat)$, then~\eqref{eq:mean_update} becomes
\alignn{
\nabla\conj(\hat \nat_{n}) 
= \nabla\conj(\hat \nat_{n-1}) - \lr_n g_n(\hat \nat_{n-1})
}
which we recognize as the update formula for stochastic mirror descent (SMD) on $f$ with mirror map $\nabla\logpart$ and step-size schedule $\lr_n$ initialized at $\nat_0$.
\emph{MAP is an instantiation of SMD.}
Note that this analogy does not hold for the MLE %($n_0=0$)
as it does not have any initialization to start from.

@@

\section{Optimization Perspective}
\label{sec:optimization}
As we saw in the technical background, MAP can be interpreted as stochastic mirror descent (SMD).
This means that \textbf{1)} we may obtain a convergence rate for MAP from an optimization analysis, and \textbf{2)} any insights gained from the MAP may inform further design and analysis of SMD.
We first review the assumptions of relative smoothness, useful to deal with non-smooth functions, before investigating 3 analysis of SMD with the MAP.
\rlp{mention that examples have so far been lacking for relative SMD.}

\subsection{Relative Smoothness}
Mirror descent (MD)\footnote{
Also known as
Bregman (proximal) gradient, relative gradient descent or NoLips.
}
\citep{nemirovski1983problem,beck2003mirror}
%Nemirovski introduced the algorithm, and Beck established the connection with Bregman divergences, e.g. the argmin form.
and SMD 
\citep{nemirovski2009robust,ghadimi2012optimal} 
are typically encountered in non-smooth (online) optimization,
under bounded (or Lipschitz) gradient assumption on the objective $f$
and strong convexity assumption on the potential $\logpart$
\citep[Th.4.2(MD) \& Th.6.3(SMD)]{bubeck2015convex}.
In our case, none of these assumptions may hold. 
For instance $\logpart = -\log$ is neither smooth nor strongly convex.
